A Large-Scale Study of the Use of Eval in R

### 2020-08-28 inst/paper/main.tex
minor

The R Project is a tool for data analysis.  At its heart is a {vectorized,
  dynamic, lazy, functional, object-oriented} programming language with a
rather unusual combination of features~\cite{ecoop12} designed to ease
learning by non-programmer and enable rapid development of statistical
methods~\cite{R96}.  The language was designed as a successor to
S~\cite{S88}.

In R, most data types are vectorized. Values are constructed by the
\c{c(...)} function: \c{c("hi","ho")} creates a vector of two strings. The
language does not differentiate scalars from vectors. To enable equational
reasoning, R copies values accessible through multiple variables when they
are written to.  Values can be tagged by user-defined attributes. For
instance, one can attach the attribute \c{dim} to the value
\c{x<-c(1,2,3,4)} by evaluating \c{attr(x,"dim")<-c(2,2)}.  Once done,
arithmetic functions will treat \c x as a \c{2x2} matrix. Another attribute
is \c{class} which can be bound to a list of names. For instance,
\c{class(x)<-"human"}, sets the class of \c{x} to \c{human}.  Attributes are
used for object-oriented dispatch.  Every linguistic construct is desugared
to a function call, even control flow statements, assignments, and
bracketing. Furthermore, all functions can be redefined in user code. This
makes R both flexible and challenging to compile~\cite{dls19}.

Arguments to user-defined functions are bundled into a thunk called a
promise. Logically, a promise combines an expression's code, its
environment, and its value.  To access the value of a promise, one must
force it. Forcing a promise triggers evaluation and the computed value is
captured for future reference.

@@

The arguments are as follows: \c{expr} is the value to be evaluated,
\c{envir} is the environment in which \c{expr} is to be evaluated, and
\c{enclos} is relevant when \c{envir} is a list or a data frame, then it
specifies where R looks for objects not found in \c{envir}. This can be
\c{NULL} (interpreted as the base package environment, \c{baseenv()}) or an
environment.

The semanitcs of \eval is to evaluate \c{expr} in the environment specified
by \c{envir} and return the computed value. If \c{envir} is not specified,
then the default is \c{parent.frame()} (the environment where the call to
\eval was made).  Values to be evaluated can be of types \c{call} (AST for
function calls) or \c{expression} (AST nodes for arbitrary expressions) or
\c{name} (when the symbol is looked up in the current scope and its binding
is evaluated), a promise or any of the basic types such as vectors,
functions and environments (which are returned unchanged).

The \c{evalq} form is equivalent to \c{eval(quote(expr), ...)}. eval
evaluates its first argument in the current scope before passing it to the
evaluator: evalq avoids this.

\c{eval.parent(expr, n)} is a shorthand for \c{eval(expr, parent.frame(n))}.

If envir is a list (such as a data frame) or pairlist, it is copied into a
temporary environment (with enclosure enclos), and the temporary environment
is used for evaluation. So if expr changes any of the components named in
the (pair)list, the changes are lost.

If envir is NULL it is interpreted as an empty list so no values could be
found in envir and look-up goes directly to enclos.

\c{local} evaluates an expression in a local environment. It is equivalent to
evalq except that its default argument creates a new, empty
environment. This is useful to create anonymous recursive functions and as a
kind of limited namespace feature since variables defined in the environment
are not visible from the outside.

The \c{expr} argument of \eval is first evaluated in the current scope, then
passed to the \eval. Quoting that argument prevents if from being evaluated
in the current environment. \c{evalq} is a shortcut for \c{eval(quote(expr),
  ...)}. In the following code snippet, variable \c{r} is first looked up in
the current scope with \eval whereas it is looked up in \c{env} with
\c{evalq}.

@@

\eval will return unchanged basic types such as vectors, functions, strings
and environments. Promises are forced by \eval. Symbols are looked up in the
current environment. Calls and expressions are evaluated.

>>>>

The R Project is a tool for data analysis.  At its heart is a {vectorized,
  dynamic, lazy, functional, object-oriented} programming language with an
unusual combination of features~\cite{ecoop12} designed to ease learning by
non-programmer and enable rapid development of statistical
methods~\cite{R96}.  The language was designed as a successor to
S~\cite{S88}.

Most R data types are vectorized. Values are constructed by the \c{c(...)}
function, e.g., \c{c("hi","ho")} creates a vector of two strings.  R does
not differentiate scalars from vectors. To enable equational reasoning, R
copies values accessible through multiple variables when they are updated.
Values can be tagged by user-defined attributes. For instance, one can
attach the attribute \c{dim} to the value \c{x<-c(1,2,3,4)} with
\c{attr(x,"dim")<-c(2,2)}.  This causes arithmetic functions to treat \c x
as a 2x2 matrix. Another attribute is \c{class} which can be bound to a list
of names, e.g., \c{class(x)<-"human"}. This sets the class of \c{x} to
\c{human}, classes are used for object-oriented dispatch.  Every R
linguistic construct is desugared to a function call, even control flow
statements, assignments, and bracketing. Furthermore, all functions can be
redefined in user code. This makes R both flexible and challenging to
compile~\cite{dls19}.

Arguments to user-defined functions are bundled into thunks called
promises. Logically, a promise combines an expression's code, its
environment, and its value.  To access the value of a promise, one must
force it. Forcing a promise triggers evaluation and the computed value is
captured for future reference.

@@

Arguments are as follows: \c{expr} is the value to be evaluated, \c{envir}
is the environment in which it is to be evaluated, and \c{enclos} is
relevant when \c{envir} is a list or a data frame, then it specifies where R
looks for objects not found in \c{envir}. This can be \c{NULL} (interpreted
as the base package environment, \c{baseenv()}) or an environment.

When called, \eval evaluates \c{expr} in the environment specified by
\c{envir} and return the computed value. If \c{envir} is missing, then the
default is \c{parent.frame()} (the environment where the call to \eval was
made).  Values to be evaluated can be of types \c{call} (AST for function
calls) or \c{expression} (AST nodes for arbitrary expressions), \c{name}
(when the symbol is looked up in the current scope and its binding is
evaluated), promise, or any of the basic types such as vectors and
environments (which are returned unchanged).

The \c{evalq} form is equivalent to \c{eval(quote(expr), ...)}.  Quoting
that argument prevents if from being evaluated in the current environment.
In the following code snippet, variable \c{r} is first looked up in
the current scope with \eval whereas it is looked up in \c{env} with
\c{evalq}.

@@

\c{eval.parent(expr, n)} is a shorthand for \c{eval(expr, parent.frame(n))}.
\c{local} evaluates an expression in a local environment. It is equivalent
to evalq except that its default argument creates a new, empty
environment. This is useful to create anonymous recursive functions and as a
kind of limited namespace feature since variables defined in the environment
are not visible from the outside.
